---
title: "Reproducible Research Week 2 Project"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
#Read the zip file
    fileName <- "repdata_data_activity.zip"
    data <- unzip(fileName)
    data <- read.csv(data)
    
#Preprocess the data to read the right type
    data$date <- as.Date(data$date)
```

## R Markdown

Below is the histogram of plots of the sample:

```{r , echo=FALSE}
    
    cumStepPerDay <- aggregate(data$steps, by=list(Category=data$date), FUN=sum)
    names(cumStepPerDay)[2] = 'Steps'
    hist(cumStepPerDay$Steps,10,
         main = paste("Histogram of" , "Cumulative Steps per Day"),
         xlab='Cumulative steps per Day')
    
```

Mean and Median number of steps.
```{r, echo=FALSE}
    paste('Mean number of steps is',mean(cumStepPerDay$Steps,na.rm=TRUE))
    paste('Median number of steps is',median(cumStepPerDay$Steps,na.rm=TRUE))
```

Plot of average number of steps for every interval.
```{r, echo=FALSE}
    fiveMinInt <- aggregate(data$steps,by=list(Category=data$interval), FUN=mean, na.rm=TRUE)
    names(fiveMinInt)[2] = 'Steps'
    
    plot(fiveMinInt$Category,fiveMinInt$Steps,type='l',
         main='Average number of steps for every Interval',
         ylab='Average number of steps',
         xlab='Five minute interval (s)')
```

Which five minute interval has the most steps?
```{r,echo=FALSE}
    i <- which(fiveMinInt$Steps == max(fiveMinInt$Steps))
    fiveMinInt[i,1]
```

How many rows have NA?
```{r, echo=FALSE}
    sum(is.na(data$steps))
```

Replace all NA values with mean of interval.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
    library(dplyr)

    head(data)

    df <- data %>% 
              group_by(interval) %>% 
                  mutate(steps = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))
    head(df,100)
```

Histogram after imputting data.


```{r, echo=FALSE}
    cumStepPerDay1 <- aggregate(df$steps, by=list(Category=df$date), FUN=sum)
    names(cumStepPerDay1)[2] = 'Steps'
    hist(cumStepPerDay1$Steps,10,
         main = paste("Histogram of" , "Cumulative Steps per Day"),
         xlab='Cumulative steps per Day')
```

Mean and Median number of steps after imputting data.
```{r, echo=FALSE}
    paste('Mean number of steps is',mean(cumStepPerDay1$Steps))
    paste('Median number of steps is',median(cumStepPerDay1$Steps,))
```
Imputting the data changes the median number of steps but not the mean.

```{r, echo=FALSE}
    data$date <- as.Date(data$date)
    day <- weekdays(data$date)
    data <- cbind(data,day)
    weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
    weekdayData <-  data[day %in% weekdays1,]
    weekendData <-  data[!(day %in% weekdays1),]
    fiveMinInt1 <- aggregate(weekdayData$steps,by=list(Category=weekdayData$interval), FUN=mean, na.rm=TRUE)
    names(fiveMinInt1)[2] = 'Steps'
    fiveMinInt2 <- aggregate(weekendData$steps,by=list(Category=weekendData$interval), FUN=mean, na.rm=TRUE)
    names(fiveMinInt2)[2] = 'Steps'
    
    par(mfrow=c(2,1),mar=c(4,6,3,3))
    plot(fiveMinInt1$Category,fiveMinInt1$Steps, type="l",ylab='Number of steps',xlab='Interval',main='Weekday')
    plot(fiveMinInt2$Category,fiveMinInt2$Steps, type="l",ylab='Number of steps',xlab='Interval',main='Weekend')
```